---
title: "Homework2"
author: "Ralph"
format: html
editor: visual
---

## Task 1:

Storing both the pre-treatment and post-treatment readings using the c() function in order to combine
the data.
```{r}
preTreatment <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

postTreatment <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Next, we assign names to these vectors by using the paste() function.

```{r}
names(preTreatment) <- paste("Subject", 1:20, sep = "_")
names(postTreatment) <- paste("Subject", 1:20, sep = "_")
```
Now we calculate the difference in the measurements by simply subtracting post treatments by pre treatments, and then calculate the average in the blood pressure change.

```{r, echo=TRUE}
differenceInTreatments <- postTreatment - preTreatment
print(differenceInTreatments)
### Mean is shown next:
print(mean(differenceInTreatments))
```
The patients who experienced a decrease in blood pressure after treatment (a positive change) are shown below using the which() function.
```{r}
positiveChanges <- which(differenceInTreatments > 0, useNames = TRUE)
print(positiveChanges)
```

### Average of those where the blood pressure decreased.
```{r}
print(mean(positiveChanges))
```

