---
title: "Homework2"
author: "Ralph"
format: html
editor: visual
---

## Task 1:

Storing both the pre-treatment and post-treatment readings using the c() function in order to combine
the data.
```{r}
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Next, we assign names to these vectors by using the paste() function.

```{r}
patients <- paste("Subject", 1:20, sep = "_")
print(patients)
```
Now we calculate the difference in the measurements by simply subtracting post treatments by pre treatments, and then calculate the average in the blood pressure change.

```{r, echo=TRUE}
diff_bp <- post_bp - pre_bp
print(diff_bp)
### Mean is shown next:
print(mean(diff_bp))
```
The patients who experienced a decrease in blood pressure after treatment (a positive change) are shown below using the which() function.
```{r}
positiveChanges <- which(diff_bp > 0, useNames = TRUE)
print(positiveChanges)
```

### Average of those where the blood pressure decreased.
```{r}
print(mean(positiveChanges))
```

## Task 2:

### Create a new data frame by assigning each of two columns the vectors post and pre treatment blood pressure readings.

```{r}
my_df <- data.frame(patients, pre_bp, post_bp, diff_bp)
my_df
```


### Returning only those rows of diff_bp where diff column is negative.

```{r}
subset(my_df, diff_bp < 0)
```

### storing TRUE in a new column on the data frame if post_bp is less than 120.

```{r}
my_df$isLessThan <- my_df$post_bp < 120
knitr::kable(my_df)
```

### Task 3:

```{r}
placebo_patients <- paste("Subject", 1:10, sep = "_")
placebo_pre_treatment <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_treatment <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo_trial_df <- data.frame(placebo_patients, placebo_pre_treatment, placebo_post_treatment)

my_list <- list(treatment = my_df, placebo = placebo_trial_df)
my_list
```

##### Next, we access the first list element using 3 different kinds of syntax. Note that the single square brackets returns both the element (data frame) and the name, whereas, double square brackets returns just the data frame.

```{r, eval=TRUE}
my_list$treatment
my_list[1]
my_list[[1]]
```

Next, In one line, we access the placebo data frame, pre_bp column.

```{r}
my_list$placebo$placebo_pre_treatment
```


### Task 4:

```{r}
my_list$treatment$status <- character(20)
my_list$placebo$status <- character(10)
my_list
```

Next, we assign statuses to each of the rows in the two data frames using control loops and if-else statements

```{r}
for(index in 1:20) {

  bp <- my_list$treatment$post_bp[index]

  if(bp <= 120) {
    my_list$treatment$status[index] <- "optimal"
  }

  if(bp > 120 & bp <= 130) {
    my_list$treatment$status[index] <- "borderline"
  }

  if(bp > 130) {
    my_list$treatment$status[index] <- "high"
  }
}

print(my_list$treatment)
```

```{r}
for(index in 1:10) {

  bp <- my_list$placebo$placebo_post_treatment[index]

  if(bp <= 120) {
    my_list$placebo$status[index] <- "optimal"
  }

  if(bp > 120 & bp <= 130) {
    my_list$placebo$status[index] <- "borderline"
  }

  if(bp > 130) {
    my_list$placebo$status[index] <- "high"
  }
}

print(my_list$placebo)
```

